---
title: Using linear regression to help on relative pricing of brazilian banks
author: Sidney Bissoli
date: '2021-07-29'
slug: linear-regression-banks
categories:
  - R
  - Finance
tags:
  - R
  - Finance
  - Linear Regression
subtitle: ''
summary: ''
authors: []
lastmod: '2021-07-29T16:24:56-03:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

In this post, I will show how to use linear regression to help on relative pricing. It is important to say that I am not dealing with absolute or intrinsic
valuation (e.g.: discounted cash flow - DCF). For the difference between pricing and valuation, see [Damodaran](https://www.youtube.com/watch?v=DeChWXTg7Og). The response variable chosen was P/B (price over book ratio), and the predictors were: 

1) Return On Equity (ROE)
2) Negotiated volume in the last 2 months
3) Net Revenues' Growth in the last 5 years
4) Equity (thousand reais)
5) If the Government is the controlling shareholder or not
6) If the companhy is a fintech or not

Data were collected from the website [Fundamentus](http://www.fundamentus.com.br/). Predictor 5 was obtained from Google. Predictor 6 was the only variable that I have built by own, from my knowledge of Brazilian banking sector. Banks were chosen because, in this sector, P/B ratio reflects more appropriately the companies' value. Brazilian banks were chosen just because I am from Brazil. So, first, let's load data and packages that we are going to use:
```{r load}
# packages needed
library(readxl)
library(GGally)
library(caret)
library(gridExtra)
library(tidyverse)
library(ggrepel)
library(sandwich)
library(jtools)

# read data
df <- read_excel("./data.xlsx", 
                 col_types = c(
                   "text","text","numeric","numeric",
                   "numeric","numeric","numeric","numeric","numeric"))
```

Let's inspect data:
```{r structure}
# view data structure
str(df)
```

"State_owned" and "Fintech" must be classified as factor variables:
```{r factor}
df$state_owned <- as.factor(df$state_owned)
df$fintech <- as.factor(df$fintech)
```

Let's visualize variables' distribution, and how they are related to each other:

```{r ggpairs, fig.dim = c(10,8)}
# visual inspection
ggpairs(df[,3:9])
```

As we can see, "price_book" is too much skewed; we will run a log transformation, to normalize it as much as we can, without losing model's interpretability.

```{r}
df$price_book_log <- log(df$price_book)
```

Moreover, 


