---
title: How to arbitrate between ITSA4 and ITUB4
author: Sidney Bissoli
date: '2021-07-31'
slug: arbitrate-itsa4-itub4
categories:
  - Finance
  - R
tags:
  - R
  - Finance
  - ITUB4
  - ITSA4
subtitle: ''
summary: ''
authors: []
lastmod: '2021-07-31T18:58:45-03:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.dim = c(10,8))
```

Itau Unibanco (ITUB4) and Itaúsa (ITSA4) are two of the most traded shares on Brazilian stock exchange. In fact, Itaúsa is a holding that manages a portfolio of companies such as: Itau Unibanco (ITUB4); Alpargatas (ALPA4); Dexco (DTEX3); Copa Energia; and Nova Transportadora do Sudeste - NTS. Despite that, considering all companies market capitalization, Itau Unibanco comprises almost 90% from Itaúsa's portfolio, such that ITUB4 and ITSA4 are highly correlated, as seen in graph below:

```{r correlation, warning = FALSE, message = FALSE}
# load packages
library(quantmod)
library(tidyverse)
library(ggpubr)

# load stocks
itsa <- getSymbols("ITSA4.SA", auto.assign = FALSE, src = "yahoo", 
                   from="2005-12-01", to="2021-06-30")
itub <- getSymbols("ITUB4.SA", auto.assign = FALSE, src = "yahoo", 
                   from="2005-12-01", to="2021-06-30")

# build one data frame
merge <- cbind(itsa[,6], itub[,6])
names(merge) <- c("itsa_price","itub_price")

# correlation ITSA4 against ITUB4
g <- ggscatter(data = merge, x = "itsa_price", y = "itub_price",
               alpha = .1, add = "reg.line", 
               add.params = list(color = "blue", fill = "lightgray"), 
               conf.int = TRUE)
g + 
  stat_cor(method = "pearson", label.x = 3, label.y = 30, size = 10) + 
  labs(x = "ITSA4",
       y = "ITUB4",
       title = "Scatterplot ITSA4 versus ITUB4") + 
  theme(axis.text=element_text(size=20),
        axis.title=element_text(size=20),
        plot.title = element_text(size=20))
```

As you can see, correlatin is equal to 0.99, which means that ITUB4 and ITSA4 mean almost the same thing. As a holding, ITSA4 is traded at a discount in relation to the sum of its investees' market capitalization. But, how much is the "fair" discount, and can we buy ITSA4 simply to get ITUB4 at lower prices? These are the questions that this post intends to answer.

Fortunatelly, Itaúsa calculates this discount for us, in a monthly basis (click [here](https://www.itausa.com.br/ListAtivos.aspx?idCanal=Afrgtrc88Ma15TQHbkasoQ==&linguagem=pt&ano=2021) to get the data). If you don't want to or don't know how to get your hands dirty, I have already done the job, which you can see on the plot below:

```{r plot discount time series}
# load packages
library(readxl)
library(scales)

# load data
df <- read_excel("./data.xlsx")

# plot time series
ggplot(data = df, aes(x = date, y = discount)) + 
  geom_line() +
  geom_hline(yintercept = mean(df$discount), 
             linetype = "dashed") + 
  geom_hline(yintercept = mean(df$discount) + sd(df$discount), 
             linetype = "dashed") + 
  geom_hline(yintercept = mean(df$discount) - sd(df$discount), 
             linetype = "dashed") + 
  scale_y_continuous(labels = percent_format(accuracy = 1)) + 
  labs(x = "Date",
       y = "Discount",
       title = "Holding's discount in relation to shares held in investee companies") + 
  theme(axis.text=element_text(size=20),
        axis.title=element_text(size=20),
        plot.title = element_text(size=20))
```

Dashed horizontal lines represent the mean (at the center), mean plus 1 standard deviation (at the top), and mean minus 1 standard deviation (at the bottom). In other words, in most part of the time, this discount ranges between something like 18% and 25%. Another way of visualizing the same information is plotting an histogram:

```{r discount histogram}
# plot distribution
ggplot(data = df, aes(x = discount)) + 
  geom_histogram(aes(y=..density..), colour="black", fill="white") + 
  geom_density(alpha=.2, fill="#FF6666") +
  scale_x_continuous(labels = percent_format(accuracy = 1)) + 
  labs(x = "Discount",
       y = "Density",
       title = "Discount's Distribution") + 
  geom_vline(xintercept = mean(df$discount), 
             linetype = "dashed", size = 1, color = "red") + 
  theme(axis.text=element_text(size=20),
        axis.title=element_text(size=20),
        plot.title = element_text(size=20))
```

Red dashed vertical line represents the mean. Distribution is rigth-skewed, which means that in some occasions the discount was too high (more than 30%). Numerically, let's view the summary stats:

```{r discount summary}
# summary statistics
summary(df$discount)
```

Mean discount is 21.5%. Minimum discount is 15.9%. 25% of observations lies between a discount of 23.4% and 30.4% (third quartile). Now we will check ITSA4 and ITUB4 cumulative returns, from Dec 2005 until now. At the end of Jul 2021, ITSA4 and ITUB4 prices were R$ 11.17 and R$ 30.30, respectively. Then, we will compare them to each other, defining success (number 1) when ITSA4 return was higher than ITUB4, and failure (number 0) when ITSA4 return was lower than or equal to ITUB4. Finally, we will inspect if there are changes in our variable of interest (discount) in both situations (success and failure). For that, first we will build our new data set:

```{r build new data set}
# get prices at the end of the month
merge <- merge[endpoints(merge, "months")]

# merge stocks and discount
data <- cbind(merge, df$discount)
names(data)[3] <- "discount"
data <- as.data.frame(data)
data <- rownames_to_column(data, "date")
data$date <- as.Date(data$date)

# add variables
data <- data %>%
  mutate(itsa_return = (11.17 / itsa_price) - 1,
         itub_return = (30.30 / itub_price) - 1,
         success = ifelse(itsa_return > itub_return, 1, 0)
         )

# print first observations
head(data, n = 10)
```

There are many ways of visualizing if there are difference on discounts on both situations. Through a boxplot:

```{r sucess failure boxplot}
# box plot of difference on returns
ggplot(data = data, 
       aes(x = factor(success), 
           y = discount, 
           fill = factor(success))) + 
  geom_boxplot() +
  scale_x_discrete(labels=c("0" = "ITSA4 <= ITUB4", "1" = "ITSA4 > ITUB4")) + 
  scale_y_continuous(labels = percent_format(accuracy = 1)) + 
  labs(x = "", 
       y = "Discount",
       title = "Returns comparison's boxplot") + 
  theme(legend.position = "none") + 
  scale_fill_manual(values = c("red3","green4")) + 
  theme(axis.text=element_text(size=20),
        axis.title=element_text(size=20),
        plot.title = element_text(size=20))
```

A density distribution:

```{r success failure density}
# density of difference on returns                           
ggplot(data = data, aes(x = discount, fill = factor(success))) +
  geom_density(alpha=.3) + 
  scale_x_continuous(labels = percent_format(accuracy = 1)) + 
  labs(x = "Discount",
       y = "Density",
       title = "Returns comparison's distribution") + 
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size=20),
        axis.text=element_text(size=20),
        axis.title=element_text(size=20),
        plot.title = element_text(size=20)) + 
  scale_fill_manual(labels = c("ITSA4 <= ITUB4", "ITSA4 > ITUB4"),
                    values = c("red","green3"))
```

Or a histogram:

```{r success failure histogram}
# histogram of difference on returns                           
data %>%
  mutate(success = recode(success,
                          "0" = "ITSA4 <= ITUB4",
                          "1" = "ITSA4 > ITUB4")) %>%
  ggplot(aes(x = discount, fill = factor(success))) +
  geom_histogram(alpha=.4, color = "black") + 
  facet_wrap(~success) + 
  scale_x_continuous(labels = percent_format(accuracy = 1)) + 
  labs(x = "Discount",
       y = "Count",
       title = "Returns comparison's distribution") + 
  theme(legend.position = "none",
        axis.text=element_text(size=20),
        axis.title=element_text(size=20),
        plot.title = element_text(size=20),
        strip.text = element_text(size = 20)) + 
  scale_fill_manual(values=c("red", "green3"))
```

On the three plots, we conclude that ITSA4 returns were higher than ITUB4 returns when ITSA4 was traded at a higher discount in relation to the sum of its investees' market capitalization (which is pretty intuitive). About the third plot, green bars are much higher just because it is more frequent that ITSA4 gets a better return than ITUB4.

Numerically, the differences on discounts between success and failure are:

```{r success failure numeric}
# numerical comparison
data %>%
  group_by(success) %>%
  summarise(minimum = min(discount),
            first_quantile = quantile(discount, .25),
            median = median(discount),
            mean = mean(discount),
            third_quantile = quantile(discount, .75),
            maximum = max(discount)) %>%
  data.frame()
```

Paradoxically, the higher discount occured when ITUB4 return was higher than ITSA4 return, but this is just an exception that proves the rule. Now we will categorizy the variable discount, so that in each categorization there is at least 30 observations. After that, we will calculate the probability of success (ITSA4 > ITUB4 returns) in each categorization:

```{r success probability}
# probability of success by discount (numerically)
library(Hmisc)
data$discount_cat <- cut2(data$discount, m = 30)
data %>%
  group_by(discount_cat) %>%
  summarise(prob = sum(success) / length(success)) %>%
  data.frame()

# view success probability by categorical discount
data %>%
  group_by(discount_cat) %>%
  summarise(prob = sum(success) / length(success)) %>%
  ggplot(aes(x = prob, y = discount_cat)) + 
  geom_point(size = 5, color="blue") + 
  scale_y_discrete(limits = rev(levels(data$discount_cat))) + 
  theme(panel.grid.major.x=element_blank(),
        panel.grid.minor.x=element_blank(),
        panel.grid.major.y=element_line(linetype="dashed", color="black"),
        axis.text.x=element_blank(),
        axis.text.y=element_text(size = 20),
        axis.ticks.x=element_blank(),
        axis.title=element_text(size=20),
        plot.title = element_text(size=20)) + 
  labs(x = "Probability",
       y = "Discount",
       title = "Probability of ITSA4 > ITUB4 Returns") + 
  geom_text(aes(label = percent(prob)), hjust = -.1, vjust = -1, size=7) + 
  xlim(c(.5,1))
```

As expected, the higher the discount, the higher the probability of ITSA4 returns exceed ITUB4 returns. But, again, paradoxically, there is a drop on success probability in the maximum range of discount (24.8% to 30.4%). We could suspect that success depends on other factors than discount - for instance, when the observations were obtained (or, in other words, the passage of time). We can check this coloring time series' discount, by success or failure:

```{r success failure time series}
# plot time series
ggplot(data = data, aes(x = date, 
                        y = discount, 
                        color = factor(success), 
                        group = 1)) + 
  geom_line(size = 1) + 
  labs(x = "Date",
       y = "Discount",
       title = "Discounts time series, by success and failure") + 
  scale_y_continuous(labels = percent_format(accuracy = 1)) + 
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size=20),
        axis.text=element_text(size=20),
        axis.title=element_text(size=20),
        plot.title = element_text(size=20)) +
  scale_color_manual(values=c("red", "green3"), 
                     labels=c("ITSA4 <= ITUB4", "ITSA4 > ITUB4"))
```

The drop on probability seen on previous graph is due to the first observations (end of year 2005 and beginning of 2006), when the pattern studied in this post didn't work very well.

At Jun 2021, ITSA4 was being traded at a discount of 24%. And then, are you going to buy, keep or sell?

Obs.: you can get data set about ITSA4 discounts and codes on my Github repository, clicking (here)[https://github.com/Sidonesio/ITSA4].
